---
title: "Hwang_AUCell_onlytumor"
output: pdf_document
date: "2024-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

First point is to load the data, get it into seurat and filter for tumor cells

```{r}
# Load sparse expression matrix
expression_matrix <- ReadMtx(
  mtx = "data/Hwang_Nature_2022/GEO/matrix.mtx",
  features = "data/Hwang_Nature_2022/GEO/features.tsv",
  cells = "data/Hwang_Nature_2022/GEO/barcodes.tsv",
  skip.cell = 0
)
metadata <- read_tsv("data/Hwang_Nature_2022/GEO/metadata.tsv")  %>% column_to_rownames("...1") 
seurat_object <- CreateSeuratObject(counts = expression_matrix)
seurat_object <- AddMetaData(seurat_object, metadata = metadata)

## Load original UMAP
## UMAP
UMAP_Coord <- read_tsv("data/Hwang_Nature_2022/GEO/obsm/X_umap.tsv", col_types = c("cnn")) %>%
  rename(UMAP_1 = "0",UMAP_2 = "1") %>% 
  column_to_rownames("...1") 


UMAP_coordinates_mat <- as(UMAP_Coord, "matrix")

seurat_object[['UMAP']] <- CreateDimReducObject(embeddings = UMAP_coordinates_mat, key = "UMAP_", global = T, assay = "RNA")


# Filter 
## keep tumor cells
seurat_object <- subset(x = seurat_object, 
                  subset = new_celltypes == "Epithelial-Malignant")



# Remove unused objects
rm(expression_matrix)
rm(metadata)
```

Then load and create the ISRact signature, dividing it in high and low

```{r, echo=FALSE}

# Build gene sets
## ISRact
pca_pdx <- read_rds("data/Classifiers/pca_pdx_ENZO.RDS")
ISRact_contributions <- sort(pca_pdx$rotation[,"PC1"])
### translate to Gene ID
ensembl75 <- useEnsembl(biomart = "genes",
                        dataset = "hsapiens_gene_ensembl",
                        version = 75)#listAttributes(ensembl75, page="feature_page")

annot_ensembl75 <- getBM(attributes = c('ensembl_gene_id',
                                        'external_gene_id'), mart = ensembl75)

translate = deframe(annot_ensembl75[c("ensembl_gene_id", "external_gene_id")])

ISRact_info <- tibble(EnsemblID = names(ISRact_contributions),
                      gene_name = translate[names(ISRact_contributions)],
                      PC1_score = ISRact_contributions,
                      class = if_else(ISRact_contributions > 0, "ISRac_high", "ISRac_low"))

ggplot(ISRact_info, aes(x = PC1_score)) +
  geom_density() +
  geom_rug(aes(color = class)) +
  theme_bw()

### extract genes
ISRact_high <- dplyr::filter(ISRact_info, class == "ISRac_high") %>%
  dplyr::select("gene_name") %>% 
  deframe()

ISRact_low <- dplyr::filter(ISRact_info, class == "ISRac_low") %>%
  dplyr::select("gene_name") %>% 
  deframe()

### Build Gene set objects
ISRactGeneSets <- c(
  GeneSet(ISRact_high, setName=paste0("ISRact_high(", length(ISRact_high),")")),
  GeneSet(ISRact_low, setName=paste0("ISRact_low(", length(ISRact_low),")")))

geneSets <- GeneSetCollection(c(ISRactGeneSets))
```
Now, Lets filter the seurat object so it only consider ISRact_high genes and standardize it

```{r, echo=FALSE}
seurat_object <- NormalizeData(seurat_object, normalization.method = "LogNormalize", scale.factor = 10000)
seurat_f <- ScaleData(seurat_object, features = c(ISRact_low, ISRact_high))

input_AUCell <- seurat_f@assays$RNA$scale.data
```

Then I run AUCell and adhere the results to Seurat object
```{r}

## Calculate ranking
cells_rankings <- AUCell_buildRankings(input_AUCell, plotStats=TRUE)

## calculate AUC score
cells_AUC <- AUCell_calcAUC(geneSets = geneSets, cells_rankings)
```
Now some exploratory plots of the AUC_scores

```{r}
AUC_results <- slot(slot(cells_AUC, "assays"), "data")$AUC %>%
  t() %>%
  as_tibble(rownames="CellID") %>%
  column_to_rownames("CellID")

  barplot(height = deframe(AUC_results))
```

```{r}
seurat_object <- AddMetaData(seurat_object, metadata = AUC_results)

# UMAP plot
AUC_names = names(AUC_results)
FeaturePlot(seurat_object, reduction = "UMAP", features = AUC_names)
```

Now Its time to check this on a specific UMAP, precomputed only on the Tumor cells

```{r}
library(harmony)
seurat_object <- FindVariableFeatures(seurat_object, selection.method = "vst", nfeatures = 2000) 
seurat_object <- ScaleData(seurat_object)

seurat_object <- RunPCA(seurat_object)
seurat_object <- RunHarmony(seurat_object,"pid")
seurat_object <- RunUMAP(seurat_object, reduction = "harmony",dims=c(1:30))

```

And the visual representation

```{r}
FeaturePlot(seurat_object, reduction = "umap", features = AUC_names)
```

Now I create a ISRAct High minus low to get the differenece between both 

```{r}
ISRact_diffAUC <- dplyr::mutate(AUC_results, ISR_ratio = get(AUC_names[1]) - get(AUC_names[2])) %>% dplyr::select(ISR_ratio)
seurat_object <- AddMetaData(seurat_object, metadata = ISRact_diffAUC)

```

Now here I create the plots I will share, starting with some parameters as the features to plot
```{r}
features = AUC_names[2] # "ISR_ratio" | AUC_names[1/2]
plot1 <- FeaturePlot(seurat_object, reduction = "umap",  features = features, ncol = 2) & 
  scale_colour_gradient2(low = 'blue', mid = 'white', high = 'red') # change reductuion to "UMAP" | "umap" 

```

And now I compute the neighbours and cluster this umap to compare with ISRact

```{r}
# calculate clusters
seurat_object <- FindNeighbors(seurat_object,reduction = "umap", dims=1:2)
seurat_object <- FindClusters(seurat_object, resolution = 0.1)

plot2 <- DimPlot(seurat_object, group.by =  "seurat_clusters", reduction = "umap", ncol = 2)
```
And at this moment I compare all these clusters with some violin plots

```{r}
library(patchwork)
plot3 <- VlnPlot(seurat_object, features = c(features), group.by = "seurat_clusters", pt.size = 0) + geom_hline(yintercept=0, linetype = "dashed")

(plot1 & NoLegend() | plot2 & NoLegend()) / plot3  + plot_layout(heights = c(2,1), guides = 'collect')
```